### 2016-06-16 Manuscript/SW_early_life_telomeres_SOM.Rmd
More edits - lifelong dynamics!

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID)+(1|TerritoryID)+(1|PlateID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

#Run the model
lm1 <- glmer(DeltaRTLF ~ Agemonths+(Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID) + (1|TerritoryID),
            data = temp,
            family = 'binomial',
            REML = FALSE)
lm1 <- standardize(lm1)

>>>>

temp <- subset(juv,!is.na(cenTQ))

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + cenTQ) + (1|BirdID)+(1|FieldPeriodID)+(1|TerritoryID),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

#Run the model
lm1 <- lmer(DeltaRTL ~ Agemonths*(Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|BirdID) + (1|FieldPeriodID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

### 2016-06-16 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More edits - lifelong dynamics!

**Figure 4** Longitudinal telomere dynamics in subset(dd,Agemonths<12)enile Seychelles warblers. **A** subset(dd,Agemonths<12)enile telomere length in relation to adult telomere length measured in the same individual. **B** $\Delta$RTL in relation to the time between sampling events. Lines and shaded areas represent fitted values and 95% confidence limits from general linear models. Raw data have been ommited from **B** for the sake of clarity, but are plotted in Figure S1.

**Figure 5** Factors affecting $\Delta$RTL in subset(dd,Agemonths<12)enile Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with $\Delta$RTL as the response variable (see main text for details; TQ = territory quality). **B** $\Delta$RTL in relation to tarsus length and sex.

@@

#Run the model
lm2 <- lmer(RTL ~ Agemonths * (Tarsus) + Helper + NonHelper + Sex + BodyMass  + TQ  + (1|FieldPeriodID) + (1|TerritoryID)+(1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

temp <- subset(dd,Agemonths<12)

@@

Fig3B <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(col = `Age (months)`,
           fill =  `Age (months)`,
           y = RTL,
           x=TQ)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 2.1,label = 'C')

@@

temp <- Loss
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig4A <- ggplot(temp,
       aes(x = RTL,
           y = RTL1,
           col = LossF)) +
  ylab('Adult RTL') + 
  xlab('subset(dd,Agemonths<12)enile RTL') +
    xlim(c(0,3))+
  ylim(c(0,3))+
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c('none'),
        legend.title = element_text(size = 12),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
  geom_abline(slope = 1,intercept = 0,lty = 2)+
  geom_point()+
  scale_colour_manual(values = c('navy','red'))+
  annotate('text',x = 0,y = 3,label = 'A')

>>>>

**Figure 4** Longitudinal telomere dynamics in juvenile Seychelles warblers. **A** juvenile telomere length in relation to adult telomere length measured in the same individual. **B** $\Delta$RTL in relation to the time between sampling events. Lines and shaded areas represent fitted values and 95% confidence limits from general linear models. Raw data have been ommited from **B** for the sake of clarity, but are plotted in Figure S1.

**Figure 5** Factors affecting $\Delta$RTL in juvenile Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with $\Delta$RTL as the response variable (see main text for details; TQ = territory quality). **B** $\Delta$RTL in relation to tarsus length and sex.

@@

temp <- subset(dd,!is.na(cenTQ))
temp$LogAge <- log(temp$Agemonths)

#Run the model
lm2 <- lmer(RTL ~  LogAge* (Tarsus + NonHelper + Sex + BodyMass  + cenTQ)  + (1|FieldPeriodID) + (1|BirdID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

@@

Fig3B <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point(col = 'grey')+
  stat_smooth(method = 'lm',col = 'darkgrey')+
  annotate('text',x = 17,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(
           y = RTL,
           x=Agemonths,
       col = factor(Helper))) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm',formula = y~log(x))+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 3,y = 2.1,label = 'C')

@@

temp <- dd3
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig4A <- ggplot(temp,
       aes(x = RTL,
           y = RTL1,
           col = LossF)) +
  ylab('Adult RTL') + 
  xlab('juvenile RTL') +
    xlim(c(0,3))+
  ylim(c(0,3))+
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c('none'),
        legend.title = element_text(size = 12),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
  geom_abline(slope = 1,intercept = 0,lty = 2)+
  geom_point()+
  scale_colour_manual(values = c('navy','red'))+
  annotate('text',x = 0,y = 3,label = 'A')
